# Reproducible Research: Peer Assessment 1    
=======================================================
8/17/2014      
Swift-AAM

- - -
### Loading and preprocessing the data
```{r load_data, results='markup'}
Data <- read.csv("E:\\activity.csv", header=T)

```
```{r show_summary, results="asis"}
library(xtable)
#Set the output options for numbers
options(scipen = 9, digits = 3 )
print(xtable(summary(Data)), type="html")
```
- - -
### What is mean total number of steps taken per day?
```{r steps_day, results='markup'}
library(plyr)
spd <- ddply(Data, .(date), summarize, steps = sum(steps))
mean.spd <- mean(spd$steps, na.rm = TRUE)
median.spd <- median(spd$steps, na.rm = TRUE)
```
The mean total number of steps per day is `r mean.spd` and the median is `r median.spd`.
- - -
**Histogram of the total number of steps taken each day**
```{r hist_steps_day, results='markup'}
hist(spd$steps, xlab="Total number of steps per day", main=NULL, col="blue")
abline(v=mean.spd, col = c("gray"))
abline(v=median.spd, col = c("black"), lty = 2)
```
- - -
### What is the average daily activity pattern?
```{r steps_interval, results='markup'}
spi <- ddply(Data, .(interval), summarize, steps = mean(steps, na.rm=TRUE))
max.spi <- max(spi$steps)
max.int <- spi[spi$steps==max(max.spi),1]
```
**Plot of the average number of steps taken each interval**
```{r plot_steps_int, results='markup'}
plot(spi$interval,spi$steps, ylab="Average number of steps per interval",xlab="Interval", main=NULL, type="l", lwd=2, col="green")
abline(h=max.spi, col = c("red"))
abline(v=max.int, col = c("black"))
```
The maximum number of steps per interval in average is `r max.spi` and corresponds to the interval `r max.int`.
- - -
### Imputing missing values
```{r calc_na, results='markup'}
miss <- dim(Data[is.na(Data$steps),])[1]
```
There are `r miss` rows with missing values.
```{r filling_na, result='markup'}
#Copy of Data
Data1 <- Data
#Filling na values with total steps means (Not good enough)
# replace(Data1$steps, is.na(Data1$steps), mean(Data1$steps, na.rm=T))
#Filling na values with interval's steps means (Much better)
row.names(spi)<- spi$interval
ind <- which(is.na(Data1$steps))
Data1[ind,1] <- spi[as.factor(Data1[ind,3]),2]
detach(Data)
attach(Data1)
```
- - -
### What is the new mean total number of steps taken per day?
```{r steps_day1, results='markup'}
spd1 <- ddply(Data1, .(date), summarize, steps = sum(steps))
mean.spd1 <- mean(spd1$steps, na.rm = TRUE)
median.spd1 <- median(spd1$steps, na.rm = TRUE)
```
The *new* mean total number of steps per day is `r mean.spd1` and the *new* median is `r median.spd1`.
```{r miss_days, results='markup'}
mdays <- dim(spd[is.na(spd$steps),])[1]
mdays1 <- dim(spd1[is.na(spd1$steps),])[1]
```
*Do these values differ from the estimates from the first part of the assignment?*
The mean of the first part is `r mean.spd` as well as the mean of the second part (`r mean.spd1`). The median of the fist part is `r median.spd` is different from the median of the second part: `r median.spd1`.
*What is the impact of imputing missing data on the estimates of the total daily number of steps?*
The first total daily number of steps estimation had `r mdays` missing values, while the second one has `r mdays1` missing values.
```{r table_comp, results='asis'}
t1 <- merge(spd,spd1, by=1)
names(t1)<- c("Date","First estimate","Second estimate")
print(xtable(t1), type="html")
```
- - -
**Histogram of the new total number of steps taken each day**
```{r hist_steps_day1, results='markup'}
hist(spd1$steps, xlab="Total number of steps per day", main=NULL, col="light blue")
abline(v=mean.spd1, col = c("red"))
abline(v=median.spd1, col = c("blue"), lty = 2)
```
- - -
### Are there differences in activity patterns between weekdays and weekends?
```{r create_weekday, results='markup'}
Data1$day <- weekdays(as.Date(date))
attach(Data1)
d <- levels(factor(day))
Data1$week <- "weekday"
Data1[day==d[1],5]<-"weekend"
Data1[day==d[6],5]<-"weekend"
Data1$week <- as.factor(Data1$week)
spiw <- ddply(Data1, .(interval,week), summarize, steps = mean(steps, na.rm=TRUE))
max.spiw <- max(spiw$steps)
max.intw <- spiw[spiw$steps==max(max.spiw),1]
```
**Plot of the average number of steps taken each interval during weekends and weekdays**
```{r plot_steps_week_int, results='markup'}
library(ggplot2)
ggplot(data=spiw, aes(x=interval, y=steps, group=week)) + geom_line(aes(color=week))+ facet_wrap(~ week, nrow=2)
```